---
title: "intro-to-ggplot2"
author: "Amanda Antonelli"
date: "12/8/2019"
output: html_document
---

## Attach packages

```{r}
library(tidyverse)
library(readxl)
library(here)
```

## Read in the data

```{r}
ca_np <- read_csv(here("R-workshop-data", "ca_np.csv"))

ci_np <- read_excel(here("R-workshop-data", "ci_np.xlsx"))
```

## Make a graph

Tell R: 
1. That we're using `ggplot`
2. What data we're using, and what variables we want to plot
3. What type of graph we're creating

```{r, eval = FALSE}
ggplot(data = THIS, aes(x = THIS, y = THIS)) +
  geom_THIS()
```

Now let's create a graph of CI visitation:

Line graph
```{r}
ggplot(data = ci_np, aes(x = year, y = visitors)) +
  geom_line()
```

Scatterplot
```{r}
ggplot(data = ci_np, aes(x = year, y = visitors)) +
  geom_point()
```

Column graph
```{r}
ggplot(data = ci_np, aes(x = year, y = visitors)) +
  geom_col()
```

Area graph
```{r}
ggplot(data = ci_np, aes(x = year, y = visitors)) +
  geom_area()
```

## Intro to customizing `ggplot` graphs

- color = (update point or line color)
- fill = (update the fill characteristic of an element)
- linetype = 
- pch = (value)
= size = 
- alpha = 

Only have aes() when we're changing an element based on a variable 

```{r}
ggplot(data = ci_np, aes(x = year, y = visitors)) + 
  geom_line(color = "limegreen", 
            linetype = "dashed")
```

```{r}
ggplot(data = ci_np, aes(x = year, y = visitors)) + 
  geom_point(color = "blue",
             pch = 17,
             size = 4, 
             alpha = 0.5)
```

```{r}
ggplot(data = ci_np, aes(x = year, y = visitors)) + 
  geom_point(color = "blue",
             pch = 4,
             size = 4, 
             alpha = 0.9)
```

I have some awesome text here, and below is my amazing finalized graph.

```{r, echo = FALSE}
ggplot(data = ci_np, aes(x = year, y = visitors)) + 
  geom_point(color = "blue",
             pch = 4,
             size = 4, 
             alpha = 0.9)
```

## Customize elements based on variable values or outcomes
```{r}
ggplot(data = ci_np, aes(x = year, y = visitors)) + 
  geom_point(aes(size = visitors,
             color = visitors), 
             pch = 17, 
             alpha = 0.5)
```

Activity: Create a column plot of CINP visitation over time, where the FILL COLOR changes based on the number of visitors, and outline color (color = ) is "magenta."

```{r}
ggplot(data = ci_np, aes(x = year, y = visitors)) + 
  geom_col(aes(fill = visitors),  
             color = "magenta")
```

```{r}
ggplot(data = ci_np, aes(x = year, y = visitors)) + 
  geom_line(size = 3, 
            color = "red") +
  theme_minimal()
```

You can also create your own themes

## Add/update labels and titles

Using `labs` 

```{r}
ggplot(data = ci_np, aes(x = year, y = visitors)) + 
  geom_point(size = 1, 
             color = "navy") +
  theme_minimal() +
  labs(x = "My x-axis title",
       y = "Annual visitors",
       title = "Channel Islands NP Visitation",
       subtitle = "(1963 - 2016)")
```

## Combine compatible geoms

Example: Add a point plot on top of a line plot:

```{r}
ggplot(data = ci_np, aes(x = year, y = visitors)) +
  geom_line(color = "purple") +
  geom_point(color = "blue",
             aes(size = year))
```

Activity: create a combined column graph plus line graph of visitation. 

```{r}
ggplot(data = ci_np, aes(x = year, y = visitors)) +
  geom_col(fill = "orange") +
  geom_line(color = "blue")
```

## Create a line graph with multiple series

```{r}

ggplot(data = ca_np, aes(x = year, y = visitors, color = park_name)) +
  geom_line()

```

## Faceting ggplot graphs

- Each series (group) gets its own plot panel

```{r}
ggplot(data = ca_np, aes(x = year, y = visitors, color = park_name)) +
  geom_line(show.legend = FALSE) +
  theme_minimal() +
  facet_wrap(~ park_name)
```

## Use ggsave to export a plot

Will export most recently created plot

```{r}
ggsave("np_graph.png", dpi = 180, width = 8, height = 7)
```

#Sync back to GitHub Remote