---
title: "Filter/Joins (like VLOOKUP in Excel)"
author: "Amanda Antonelli"
date: "12/8/2019"
output: html_document
---

## Attach packages

tidyverse, readxl, here

```{r, include = FALSE}
library(tidyverse)
library(readxl)
library(here)
```

Read in 3 files:

- invert_counts_curated.xlsx
- fish_counts_curated.csv
- kelp_counts_curated.xlsx

```{r}
invert_counts <- read_excel(here("R-workshop-data/invert_counts_curated.xlsx"))
  
fish_counts <- read_csv(here("R-workshop-data/fish_counts_curated.csv"))
  
kelp_counts_abur <- read_excel(here("R-workshop-data/kelp_counts_curated.xlsx"))
```

## Filter to match rows based on a character string

Only keep observations from fish counts where the common_name variable matches the string "garibaldi"

When asking, does this match? Use == 
```{r}
fish_garibaldi <- fish_counts %>% 
  filter(common_name == "garibaldi")
```

## Filter to keep or exclude observations based on numeric conditions

Create a subset where the total counts are greater than or equal to 50
```{r}
fish_over50 <- fish_counts %>% 
  filter(total_count >= 50)
```

# Activity: Create a subset that only contains observations for fish at Mohawk Reef ("mohk")
```{r}
fish_mohk <- fish_counts %>% 
  filter(site == "mohk")
```

## Filter to keep rows that match THIS *or* THAT

Example: We want to keep any observations for garibaldi, blacksmith, or black surfperch

Use %in% operator to look for any matches within a vector
```{r}
fish_3sp <- fish_counts %>% 
  filter(common_name %in% c("garibaldi", "blacksmith", "black surfperch"))

# Writing OR statements explicitly
# Use vertical line above the Return key | to indicate "or"

fish_3sp_v2 <- fish_counts %>% 
  filter(common_name == "garibaldi" | common_name == "blacksmith" | common_name == "black surfperch")
```

## Filter to keep rows that match THIS *and* THAT (multiple variables)

I want to create a subset called urchin_abur_mohk that only keeps observations where:

- site is "abur" AND
- species is "purple urchin"

Some ways to do this:

- separate "AND" conditions within a single filter line by a comma
- separate "AND" conditions within a single filter line by an ampersand (&)
- putting each filter condition within a separate filter line

```{r}
urchin_abur_mohk <- invert_counts %>% 
  filter(site %in% c("abur", "mohk") & common_name == "purple urchin")

```

